To Airbnb?: Factors impacting short-term leasing preference

### 2018-11-23 papers/full_paper.Rmd
Update to literature review.

Existing research into short-term rental markets is sparse. Within the existing work, the focus remains centered on rural and/or amenity-based STRs such as beach-front and ski-resort access properties.  This focus results from the fact that, until a few years ago, the STR market itself operated almost wholly within these market segments. The shift to small, urban dwellings within the STR market is relatively new and explains the dearth of past work in this direction. With the rapid growth of Airbnb, and the controversies surrounding it, interest in the short-term rental market has increased among researchers as well as investors and industry analysts.

Much of the existing work focuses on price or rent formation in amenity-driven second home markets.  Specifically, the impacts of amenities such as water access [@nelson2010valuing], water quality [@gibbs2002hedonic; @clapper2014water], proximity to ski facilities  [@soguel2008impact; @nelson2010valuing] and theme parks [@tsai2015impact] are shown to positively impact sales prices in tourism-dominated areas.  Host policies can also matter as @benjamin2001value find that smoking prohibitions influence the weekly rental rates for vacation homes.

Focusing specifically on the short-term rental market, @cassidy2007tourist suggest that within Australia's Gold Coast the STR industry lacks sophistication and efficiency.  Most notably, they find that most nightly or weekly rates are set by resident unit managers and are often under-priced as there is little accountability to owners.  The authors also note that very little research has been put toward understanding this growing market segment.

A collection of recent work has looked into the impact of short-term markets on housing prices and affordability. This research examines the influence of Airbnb on existing housing prices nationally [@barron2018sharing] as well as focused work in New York [@sheppard2016airbnb], Los Angeles [@lee2016airbnb] and Boston [@merante2017home]. Though very broad analyses, all suggest that Airbnb has had a positive impact on local home prices and/or a negative impact on affordability. Other research examines the impact of short-term rentals on the hotel market, with @gutierrez2016airbnb finding that Airbnb and hotels have similar location patterns in Barcelona and @zervas2014rise showing an inverse relationship between Airbnb listings and hotel revenues.

@@

@@

Overall, the extant literature on short-term rentals is limited, but growing.  Of the three types of research -- pricing, impacts on housing market and legalities/externalities -- none specifically address the decision faced by an owner who is trying to decide between a long- or short-term leasing strategy. The closest analog in the literature is a piece by @larson2009pur examining the choice between purchasing a time-share or simply paying for hotel rooms.  Theoretically, the literature shows that in most cases there is a demonstratable difference between first and second home market dynamics due to the use differences of the occupants.  We hypothesize that this same thinking can be extended to the difference between short-term rental occupants (tourists and business travelers) and long-term rental occupants (local workers and retirees) and, as a result, there should be differences in the market which make certain housing products and locations preferable (more profitable) under a short-term versus long-term leasing strategy.

>>>>

Existing research into short-term rental markets is sparse. Within the existing work, the focus remains centered on rural and/or amenity-based STRs such as beach-front and ski-resort access properties.  This focus results from the fact that, until a few years ago, the STR market itself operated almost wholly within these market segments. The shift to small, urban dwellings within the STR market is relatively new and explains the dearth of past work in this direction. With the rapid growth of Airbnb, and the controversies surrounding it, interest in the short-term rental market has increased among researchers as well as investors and industry analysts [@manning2018hotel].

Much of the existing work focuses on price or rent formation in amenity-driven second home markets.  Specifically, the impacts of amenities such as water access [@nelson2010valuing], water quality [@gibbs2002hedonic; @clapper2014water], proximity to ski facilities  [@soguel2008impact; @nelson2010valuing] and theme parks [@tsai2015impact] are shown to positively impact sales prices in tourism-dominated areas.  Host policies can also matter as @benjamin2001value find that smoking prohibitions influence the weekly rental rates for vacation homes. More recent work by Gibbs and co-authors [-@gibbs2018pricing] examine short-term rental pricing formation in large, urban areas throughout Canada.  Much like decades of work on house prices [@sirmans2006value], they find that location and physical characteristics are the primary drivers of short term rental price differentials.  Additionally, host policies are shown to play a significant role as well in these Canadian markets.

@@

Looking from the supply-side, many short-term rental markets are characteristized by a highly bifurcated set of owners and hosts. On one side are the small-time owners and hosts who manage a single or a few short-term properties.  They are constrasted by "professionals", those who have have dozens or more listings [@gibbs2018use].  Professionals are more likely to manage the more expensive, better performing and better located listings.  Additionally, they are much more likely to use dynamic pricing models [@oskam2018s; @gibbs2018use] than simply setting a price and leaving it for months at a time, regardless of daily changes in demand.

Not all "professionals", however, appear to be using the most efficient or economically sound pricing methods.  Having a vested interest in the occupancy rate matters as well. Within Australia's Gold Coast work by @cassidy2007tourist suggest that the STR industry lacks sophistication and efficiency.  Most notably, they find that most nightly or weekly rates are set by resident unit managers and are often under-priced as there is little accountability to owners.  The authors also note that very little research has been put toward understanding how property managers influence this growing market segment.

#### Externalities

The growth of the short-term market is not only a concern of profit-minded investors; the externalities, real or perceived, that have come with have gathered significant attention.  A collection of recent work has looked into the impact of short-term markets on housing prices and affordability. This research examines the influence of Airbnb on existing housing prices nationally [@barron2018sharing] as well as focused work in New York [@sheppard2016airbnb], Los Angeles [@lee2016airbnb] and Boston [@merante2017home]. Though very broad analyses, all suggest that Airbnb has had a positive impact on local home prices and/or a negative impact on affordability. Other research examines the impact of short-term rentals on the hotel market, with @gutierrez2016airbnb finding that Airbnb and hotels have similar location patterns in Barcelona and @zervas2014rise showing an inverse relationship between Airbnb listings and hotel revenues.

@@

Finally, Airbnb's unique and highly transparent peer-to-peer leasing platform has produced some concerns over equity and discrimination [@cheng2018sharing]. While official Airbnb policy addresses potential discrimination by hosts, the behaviors by guests in choosing lodging may be influencing pricing and hosting decisions as evidenced by an analysis of the San Francisco market [@kakar2018visible].  With its substantial growth and significant share of the overall urban short-term market Airbnb as a company will be the likely target of many nuisance and externality concerns in the near future.

#### Summary and Motivation

Overall, the extant literature on short-term rentals is limited, but growing.  Of the three types of research -- pricing, housing market impacts and legalities/externalities -- none specifically address the decision faced by an owner who is trying to decide between a long- or short-term leasing strategy.  The closest analog in the literature is a piece by @larson2009pur examining the choice between purchasing a time-share or simply paying for hotel rooms.

Two key findings from the existing work motivate this study. First, short-term market pricing is influenced by the same characteristics as for sale and long-term rental housing -- namely physical characteristics and location -- as well as by hosting policies [@gibbs2018pricing]. Second, there is a demonstratable difference between first and second home market dynamics due to the use differences of the occupants.  We hypothesize that these findings can be extended to investigate the differences between short-term rental occupants (tourists and business travelers) and long-term rental occupants (local workers and retirees) and, as a result, there should be differences in the market which make certain housing products and locations preferable (more profitable) under a short-term versus long-term leasing strategy.

### 2018-11-23 papers/round_2_reviews.Rmd
Update to literature review.

---
title: "To Airbnb?: Factors impacting short-term leasing preference"
subtitle: "Round 2 Reviews"
author: "Andy Krause and Gideon Aschwanden"
date: "11/8/2018"
output: pdf_document
---

>>>>

---
title: "To Airbnb?: Factors impacting short-term leasing preference"
subtitle: 'Reviewer Responses'
date: "11/8/2018"
output: pdf_document
---
